### 2016-09-25 Paper/paper.tex

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.

\usepackage{pseudocode}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsfonts}

\newcommand{\cL}{{\cal L}}

\newcommand{\func}[1]{{\sc #1}}
\newcommand{\farray}{\text{sequence}}
\newcommand{\tfunc}[1]{{\text{\func{#1}}}}
\newcommand{\val}{\func{val}}
\newcommand{\tval}{\text{\func{val}}}
\newcommand{\abs}{\func{abs}}
\newcommand{\new}{\func{new}}
\newcommand{\cost}{\func{cost}}
\newcommand{\tabulate}{\func{tabulate}}
\newcommand{\get}{\func{get}}
\newcommand{\set}{\func{set}}
\newcommand{\push}{\func{push}}
\newcommand{\size}{\func{size}}
\newcommand{\tget}{\text{\get}}
\newcommand{\tset}{\text{\set}}
\newcommand{\tpush}{\text{\push}}
\newcommand{\tsize}{\text{\size}}
\newcommand{\tnew}{\text{\new}}
\newcommand{\ttab}{\text{\tabulate}}
\newcommand{\MutableArray}{\text{MutableArray }}
\newcommand{\ArrayData}{\text{ArrayData }}
\newcommand{\Array}{\text{Array }}
\newcommand{\intref}{\text{int ref }}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

% \titlebanner{banner above paper title}        % These are ignored unless
% \preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Parallel Functional Arrays}

% \authorinfo{}{}{\\[-3.0ex]}
\authorinfo{Ananya Kumar}
           {Carnegie Mellon University}
           {ananyak@andrew.cmu.edu\\[-5.0ex]}
\authorinfo{Guy E. Blelloch}
           {Carnegie Mellon University}
           {blelloch@cs.cmu.edu}
\authorinfo{Robert Harper}
           {Carnegie Mellon University}
           {rwh@cs.cmu.edu}
\maketitle

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\begin{abstract}
The goal of our paper is to come up with functional arrays (sequences) that are as efficient as imperative arrays, can be used in parallel, and have well defined cost-semantics. Our approach is to consider arrays with functional value semantics but non functional cost semantics. Because the value semantics is functional, ``updating" an array returns a new array. We allow operations on ``older" arrays (called interior arrays) to be more expensive than operations on the ``most recent" arrays (called leaf arrays) .

We embed sequences in a language that supports fork join parallelism. Because we support fork-join parallelism, the cost of programs is non-deterministic and depends on the order in which operations are interleaved. This is because the ordering of operations can affect whether a sequence is a leaf or interior. Consequently the concurrent costs of such programs are difficult to analyze. Our main result is the derivation of a deterministic evaluational dynamics which makes analyzing the costs much easier. Our theorems are not array-specific and can be applied to any data type with different costs for operating on interior and leaf versions.

We give a wait-free concurrent sequence implementation which requires constant work for accessing and updating leaf arrays and bounded work for operations on interior arrays. We sketch out our proof of correctness for the array implementation. The key advantages of our sequence approach compared to current approaches is that our implementation requires no changes to existing programming languages, supports nested parallelism, and has well defined cost semantics. At the same time, it allows for functional implementations of algorithms like depth-first search with the same asymptotic complexity as imperative implementations.
\end{abstract}

\category{D.3.1}{Programming Languages}{Formal Definitions and Theory}
\category{F.3.2}{Logics and Meanings of Programs}{Semantics of Programming Languages}

\keywords
cost semantics, concurrency, parallel, functional data structures, persistence, arrays

\input{intro}

\input{dynamics}

\input{structural}

\input{evaluational}

\input{costproof}

\input{implementations}

\input{experiments}

\input{conclusion}

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\bibliography{references}

\end{document}